Probabilistic Catalogs from Crowded Star Fields

### 2012-07-23 paper/ms.tex
Phase transitionf igure
Broken pareto

\subsection{The Unknown Parameters}
Our model makes the following assumptions. There are an unknown number of stars
$N$ in the field. Each star has an unknown
position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
describe the {\it spatial distribution} of the stars by parameters
$\alpha$, and the distribution of fluxes (commonly known as the {\it luminosity
function}) by parameters $\beta$. In summary, the unknown parameters are:
\begin{eqnarray}
\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
\left\{f_i\right\}_{i=1}^N\right\}
\end{eqnarray}
The prior probability distribution for the unknown parameters can be factorized
using the product rule. With a variety of independence assumptions, the prior
can be written as:
\begin{eqnarray}
p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha)
p(f_i | \beta) 
\end{eqnarray}
Here, we have assumed that the luminosity function does not depend on position.
We have also assumed independence of $N$, $\alpha$ and $\beta$, which means that
learning the value of one would not tell us anything about the values of the
others. Finally, the positions and fluxes of the stars come independently from
a common distribution. If we knew the spatial distribution and the luminosity
function of the stars, then the location and flux of a particular star would
not tell us anything about the location and flux of another star.

@@

@@

\subsection{Improving the Efficiency of Reversible Jump}
Instead of parameterising by $N$ and the positions and fluxes, what if we have
$3N_{\rm max}$ latent variables $\theta$ that always exist, and transform these
into
positions and fluxes by a transformation $f$:
\begin{eqnarray}
\{\mathbf{x}, \mathbf{y}, \mathbf{f}\} &=& f(\theta)
\end{eqnarray}
The transformation can be parameterized by various parameters of its own, such
that a known fraction of the fluxes end up being nonzero, etc. This could have
various advantages, most notably: when new stars are added, they are added with
low fluxes. It also turns the variable dimension problem into a fixed dimension
one, so emcee, MultiNest or similar will now work, although the multimodality problem will
still exist.

@@

>>>>

\subsection{The Hypothesis Space and the Prior}
Our model makes the following assumptions. There are an unknown number of stars
$N$ in the field. Each star has an unknown
position $(x,y)$ in the plane of the sky, and an unknown flux $f$. We also
describe the {\it spatial distribution} of the stars by parameters
$\alpha$, and the distribution of fluxes (commonly known as the {\it luminosity
function}) by parameters $\beta$. In summary, the unknown parameters are:
\begin{eqnarray}
\theta = \left\{N, \alpha, \beta, \left\{x_i, y_i\right\}_{i=1}^N, 
\left\{f_i\right\}_{i=1}^N\right\}
\end{eqnarray}
The prior probability distribution for the unknown parameters can be factorized
using the product rule. With a variety of independence assumptions, the prior
can be written as:
\begin{eqnarray}
p(\theta) = p(N)p(\alpha)p(\beta)\prod_{i=1}^N p(x_i, y_i | \alpha)
p(f_i | \beta) 
\end{eqnarray}
Here, we have assumed that the luminosity function does not depend on position.
We have also assumed independence of $N$, $\alpha$ and $\beta$, which means that
learning the value of one would not tell us anything about the values of the
others. Finally, the positions and fluxes of the stars come independently from
a common distribution. If we knew the spatial distribution and the luminosity
function of the stars, then the location and flux of a particular star would
not tell us anything about the location and flux of another star. Really, this
is just a way of implementing exchangeability.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{Figures/likelihood.eps}
\end{center}
\end{figure}

@@

@@

\section{Broken Pareto Distribution}
The Broken Pareto distribution is:
\begin{eqnarray}
p(x | \theta) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < x_0 \\
x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
x^{-\alpha_2 - 1}, & & x > x_1
\end{array}
\right.
\end{eqnarray}
We will need the fully normalised version. Putting in two normalising constants:
\begin{eqnarray}
p(x | \theta) &=&
\left\{
\begin{array}{lcr}
0, & & x < x_0 \\
C_1x^{-\alpha_1 - 1}, & & x_0 \leq x \leq x_1 \\
C_2x^{-\alpha_2 - 1}, & & x > x_1
\end{array}
\right.
\end{eqnarray}
The normalising constants can be obtained by enforcing the following
constraints. The first constraint is that the total probability is unity:
\begin{eqnarray}
\int p(x|\theta) \, dx = 1
\end{eqnarray}
The second constraint is that the PDF should be continuous at $x=x_1$:
\begin{eqnarray}
C_1 x_1^{-\alpha_1 - 1} = C_2 x_1^{-\alpha_2 - 1}
\end{eqnarray}
